---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# darklyplot

<!-- badges: start -->
<!-- badges: end -->

The goal of darklyplot is to create simple time series plots with a dark background.  The miniminum and maximum values are highlghted, and color coded along with the y axis and x axis labels.

## Installation

You can install the released version of darklyplot from [github(https://CRAN.R-project.org) with:


And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("lenkiefer/darklyplot")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example,message=FALSE,warning=FALSE,fig.width=9}
library(darklyplot)
darklyplot(mtg_rate,"rate",labelx="roundx",n.decimals=3)
```

