---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# darklyplot

<!-- badges: start -->
<!-- badges: end -->

The goal of darklyplot is to create simple time series plots with a dark background.  The miniminum and maximum values are highlghted, and color coded along with the y axis and x axis labels.

## Installation

You can install the development version of darklyplot from [GitHub](https://github.com/) with: 

``` r
# install.packages("devtools")
devtools::install_github("lenkiefer/darklyplot")
```
## Examples

This is a basic example which shows you how to solve a common problem:

```{r example,message=FALSE,warning=FALSE,out.width="50%",fig.width=12,fig.height=8}
library(darklyplot)
darklyplot(df=mtg_rate,column="rate",labelx="roundx",n.decimals=3)
```
You can vary the chart look with several parameters:

```{r example2,message=FALSE,warning=FALSE,out.width="50%",fig.width=12,fig.heigh=8}
darklyplot(df=mtg_rate,
          column="rate",
          col="white",  #can use R color names or hex 
          n.decimals=0,
          refline=TRUE,
          refValue=5,
          refCol="purple",
          shade=TRUE,
          shadeCol="#fe5305",
          shadeAlpha=0.35,
          minCol="blue",
          maxCol="red",
          firstCol="orange",
          lastCol="pink",
          labelx="round",
          Ndodge=2)

```
Will work with any dataframe with a numberic column and a date index:

```{r example-3,message=FALSE,warning=FALSE,out.width="50%",fig.width=12,fig.height=8}
set.seed(20200704)
df_test=data.frame(date=seq.Date(from=as.Date("2020-01-01"),to=as.Date("2020-06-30"),by="1 day"))
df_test$random_variable=rnorm(NROW(df_test),0,1)
darklyplot(df=df_test,column="random_variable",n.decimals=2)+labs(title="A Random Variable")
```
